---
title: "Project 1"
author: "Matt Fein, Jeffrey Jou, Polly McKim, Elise Roberts"
date: "October 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
require(tidyr)
require(ggplot2) # Package for data visualization
require(pander) # Package to make 
require(reshape2) # Additional package for data manipulation
require(plyr) # Additional package for data manipulation
require(dplyr) # Additional package for data manipulation

outlierKD <- function(dt, var, rmv=NULL) { 
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     sd1 <- sd(var_name,na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     #
     if(is.null(rmv)) { 
       response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ") 
     } else {
       if (rmv=='y'|rmv=='yes'|rmv=='Y'|rmv=='Yes'|rmv=='YES'|rmv==TRUE ) { response = 'y' } else { response = 'n' }
     }
     #
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          cat("Nothing changed", "n")
          return(invisible(var_name))
     }
}
```

# Chapter 1: Introduction 

[insert introduction here]

# Chapter 2: Description of Data
## 2.1 Source Data

[insert description of AirBnB data]

```{r load_data, echo=F}
listing <- read.csv("listings.csv")
str(listing)
```

[describe structure of listing dataset]

# Chapter 3: Exploratory Data Analysis
## 3.1 Do the number of reviews differ by room type?

```{r remove_outliers, include=F}
outlierKD(listing, price, 'y')
outlierKD(listing, number_of_reviews, 'y')
```

> outlierKD(listing, price)  
> outlierKD(listing, number_of_reviews)

First, we remove any outliers in the number of reviews variable of the dataset using the outlierKD function courtesy of Dr. Lo. Then we explore the data through basic charts, starting with a boxplot of the number of reviews grouped by room type. The variable "room_type" is a factor variable with 3 levels: Entire home/apt, private room, and shared room. 

``` {r room_to_reviews_box, echo=F}
# Plot boxplot of number of reviews by room type
room_to_reviews <- ggplot(listing, aes(x = as.factor(room_type), y = number_of_reviews, color = room_type)) +
  geom_boxplot() + 
  theme_bw() + 
  labs(x = "Room Type",
       y = "Number of Reviews",
       color = "Room Type") 
print(room_to_reviews + ggtitle("Boxplot of Number of Reviews by Room Type")) # Print boxplot with title
```

[insert description of boxplot]

``` {r ANOVA_room_review, echo =F}
# ANOVA test to see if mean number of reviews is different between room types
# Conclusion: statistically significant, reject null hypothesis
anova_room_review <- aov(number_of_reviews ~ as.factor(room_type), data = listing)
pander(anova_room_review)

# Tukey HSD test to determine which of the room types are pairwise considered different
# Conclusion: Private to entire home  different
# Conclusion: Shared to entire home and shared to private room not different
tukey_room_review <- TukeyHSD(anova_room_review)
tukey_room_review
```

[explain ANOVA and Tukey test results/conclusions]

## 3.2 Does the price of a room differ by room type?

To begin, we start with another basic visualization of price by room type using another boxplot

``` {r price_to_room_type, echo=F, warning=F}
# Boxplot of price by room type
price_to_room_type <- ggplot(listing, aes(x = factor(room_type), y = price, color = room_type)) +
  geom_boxplot() + 
  theme_bw() +
  labs(x = "Room Type",
       y = "Price",
       color = "Room Type")
print(price_to_room_type + ggtitle("Boxplot of Price by Room Type")) # Print boxpot with title
```

[description of boxplot]

``` {r ANOVA_room_price, echo=F}
# One-way ANOVA test to see if mean prices are different between room types
# Conclusion: statistically significant, reject null hypothesis
anova_room_price <- aov(price ~ as.factor(room_type), data = listing)
pander(anova_room_price) # Print formatted table

# Tukey HSD test to determine which of the room types are pairwise considered different
# Conclusion: all different from one another 
tukey_room_price <- TukeyHSD(anova_room_price)
tukey_room_price
```

(describe ANOVA and Tukey test results/conclusions)

## 3.3 How do prices differ by neighborhood?

``` {r bar_clean, include=F}
# Separate neighborhoods by unit id
neighborhood <- 
  as.data.frame(listing) %>%
  select(id, price, neighbourhood)

neighborhood <- separate(neighborhood, neighbourhood, into = c("N1", "N2", "N3"), sep = ",")
neighborhood <-melt(neighborhood, 
       id.vars = c("id", "price"),
       variable.name = "neighborhood",
       na.rm=T) %>%
  arrange(id) 
names(neighborhood) <- c("id", "price", "number", "neighborhood")
```

``` {r freq_table, echo=F}
# Histogram of number of AirBnBs by neighborhood
freq_table <- 
  neighborhood %>%
  group_by(neighborhood) %>%
  count() %>%
  arrange(-n)

top_15 <- subset(freq_table, n > quantile(freq_table$n, 0.85))
```

``` {r neigh_bar, echo=F}
neigh_bar <- ggplot(top_15, aes(x = neighborhood, y = n)) +
  geom_bar(aes(fill=neighborhood), stat="identity") + 
  theme_bw() + 
  labs(x = "Neighborhood",
       y = "Number of AirBnbs",
       fill = "Neighborhood") + 
  theme(axis.ticks.x = element_blank()) + 
  scale_x_discrete(labels = c("Blmn.", 
                              "K St.",
                              "Kingman",
                              "Lincoln",
                              "Logan",
                              "Mt. Pleasant",
                              "Pleasant",
                              "Stanton",
                              "Truxton",
                              "Hill",
                              "Columbia Hgts",
                              "Dupont",
                              "Edgewood",
                              "Shaw",
                              "Union Station"))
print(neigh_bar + ggtitle("Top 15 Most Popular AirBnB Neighborhoods")) # Print bar chart
```

[describe bar chart]

``` {r neighborhood_scatter, echo=F}
# Scatterplot of mean price by neighborhood
price_table <- na.omit(neighborhood)
price_table <-
  price_table %>%
  group_by(neighborhood) %>%
  mutate(mean_price = mean(price, na.rm=T))

price_table <- match_df(price_table, top_15, on = "neighborhood")

price_to_neigh <- ggplot(price_table, aes(x = neighborhood, y = mean_price)) +
  geom_point(aes(color = neighborhood)) +
  theme_bw() + 
  labs(x = "Neighborhood",
       y = "Mean Price",
       color = "Neighborhood") + 
  scale_x_discrete(labels = abbreviate)
print(price_to_neigh + ggtitle("Average Price of Top 15 Most Popular Neighborhoods"))
```

[describe scatterplot]

``` {r price_to_neigh_box, echo=F}
# Boxplot of price by neighborhood
price_to_neigh_box <- ggplot(price_table, aes(x = neighborhood, y = price, color = neighborhood)) + 
  geom_boxplot() + 
  theme_bw() +
  labs(x = "Neighborhood",
       y = "Price",
       color = "Neighborhood") +
  scale_x_discrete(labels = abbreviate)
print(price_to_neigh_box + ggtitle("Boxplot of Price by Top 15 Most Popular Neighborhoods"))
```

[describe boxplot]

``` {r ANOVA_price_neigh, echo=F}
# One-way ANOVA test to see if mean prices are different among neighborhoods
# Conclusion: Statistically significant, reject the null hypothesis
anova_price_neigh <- aov(price ~ as.factor(neighborhood), data = price_table)
pander(anova_price_neigh)

# Tukey HSD test to see which neighborhoods are pairwise different
# Lots of stuff
tukey_price_neigh <- TukeyHSD(anova_price_neigh)
tukey_price_neigh
```

[describe ANOVA and Tukey test results/conclusions]

## 3.5 How does the number of reviews affect price?

``` {r price_to_reviews, echo=F}
# Bin the number of reviews by 25 review bins
listing$bins <- cut(listing$number_of_reviews, breaks = seq(0, 120, 10))
listing <- subset(listing, number_of_reviews > 0) 
listing <- 
  listing %>%
  group_by(bins) %>%
  mutate(mean_price_bins = mean(price, na.rm=T)) # Take mean price by review bin

# Scatterplot of average price (by bin) vs. number of reviews
price_to_reviews <- ggplot(listing, aes(x = as.factor(bins), y = mean_price_bins, color = mean_price_bins)) + 
  geom_point() +
  theme_bw() +
  labs(x = "Number of Reviews",
       y = "Mean Price (by bin)",
       color = "Price") + 
  theme(axis.text.x = element_text(size = 7)) 
print(price_to_reviews + ggtitle("Mean Price (by bin) vs. Number of Reviews")) # Print scatterplot with title
```

[describe scatterplot of average price (by bin) vs. number of reviews]

``` {r price_to_review_to_room, echo=F, warning=F}
# Scatterplot of price by number of reviews by room type
price_to_reviews_to_room <- ggplot(listing, aes(x = number_of_reviews, y = price, color = room_type)) +
  geom_point() + 
  theme_bw() +
  labs(x = "Number of Reviews",
       y = "Price",
       color = "Room Type")
print(price_to_reviews_to_room + facet_grid(.~room_type) + ggtitle("Scatterplot of Price by Number of Reviews")) # Print scatterplot with title
```

[describe scatterplot of average price vs. number of reviews by room type]